### 2025-04-22 cards/paper.tex
paper: add logarithms, polynomial identities and linear coefficient

\documentclass{article}
\usepackage{amsmath,amssymb,amsthm,textcomp}

\title{On Math}
\author{Jacek Walczak}
\date{February - March 2025}

\begin{document}

\maketitle

\section{Default symbols}

A - area

\section{Trigonometry}

\begin{equation}
  \begin{gathered}
    \sin\theta = \frac{\text{opposite}}{\text{hypotenuse}} \\
    \cos\theta = \frac{\text{adjacent}}{\text{hypotenuse}} \\
    \tan\theta = \frac{\text{opposite}}{\text{adjacent}} \\
    \csc\theta = \frac{1}{\sin{\theta}} =
    \frac{\text{hypotenuse}}{\text{opposite}} \\
    \sec\theta = \frac{1}{\cos{\theta}} =
    \frac{\text{hypotenuse}}{\text{adjacent}} \\
    \cot\theta = \frac{1}{\tan{\theta}} =
    \frac{\text{adjacent}}{\text{opposite}} \\
  \end{gathered}
\end{equation}

\subsection{Sine values}
\begin{equation}
  \begin{gathered}
    \sin(0) = 0 \\
    \sin(30) = \frac{1}{2} \\
    \sin(45) = \frac{\sqrt{2}}{2} \\
    \sin(60) = \frac{\sqrt{3}}{2} \\
    \sin(90) = 1
  \end{gathered}
\end{equation}

\subsection{Cosine values}
\begin{equation}
  \begin{gathered}
    \cos(0) = 1 \\
    \cos(30) = \frac{\sqrt{3}}{2} \\
    \cos(45) = \frac{\sqrt{2}}{2} \\
    \cos(60) = \frac{1}{2} \\
    \cos(90) = 0
  \end{gathered}
\end{equation}

% Tangent values

\subsection{Tangent values}
\begin{equation}
  \begin{gathered}
    \tan(0) = 0 \\
    \tan(30) = \frac{1}{\sqrt{3}} = \frac{\sqrt{3}}{3} \\
    \tan(45) = 1 \\
    \tan(60) = \sqrt{3} \\
    \tan(90) = \text{undefined}
  \end{gathered}
\end{equation}

\subsection{Formula of a general triangle}

\begin{equation}
  A = \frac{1}{2}ab \sin{C}
\end{equation}
where \textit{C} is the measure of the angle between sides \textit{a} and \textit{b} (also called \textit{included angle})

\section{Complex numbers}
\subsection{Argument of a complex number}
There is natural angle $\theta$ that \textit{z} makes with real x-axis. This angle is called \textit{the argument of z}. The numeric value of $arg(z)$ is given in radians and also
\begin{equation}
0 \leq arg(z) \leq 2\pi
\end{equation}

Example:
\begin{equation}
  \begin{gathered}
  z = 3 + 4i \\
  \tan \theta = \frac{4}{3} \\
  \theta = \arctan\frac{4}{3} \approx 0.93
  \end{gathered}
\end{equation}

\section{Geometry}
\subsection{Rigid motions}
Or sometimes - any transformations that does not change the distances
and angles between points in a figure. They are ANGLE-PRESERVING and
DISTANCE-PRESERVING.

Rigid motions types:
- translations
- rotations
- reflections

Not a rigid motions:
- dilations
- stretches

\section{Functions}
\subsection{Quadratic formula}
\begin{equation}
  x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}
\end{equation}

\subsection{Factor theorem}
For polynomial, if we know root, we can determine a factor of the
polynomial. If $p(x)$ is a polynomial and $x=r$ is a root of $p(x)$ then
\begin{equation}
  \begin{gathered}
  p(r) = 0 \\
  so (x-r) is a factor of p(x) \\
  generally: (ax-b) is a factor of a polynomial p(x) then \\
  x = \frac{b}{a} is a root of p(x)
\end{gathered}
\end{equation}

\subsection{Inverse functions}
Inverse functions are denoted by $f^{-1}(x)$.

For $f(x)$ that maps $x=2$ to $y=0$ an inverse function $f^{-1}(x)$ instantiation would be
$f^{-1}(0)=0$.

Generally it is a simple composition:
\begin{equation}
  \begin{gathered}
    f^{-1}(f(x)) = x \\
    so \\
    (f^{-1} \circ f)(x) = x
  \end{gathered}
\end{equation}

Calculate inverse function
1. swap $x$ and $y$
2. solve for $y$

Example:
\begin{equation}
  \begin{gathered}
    f(x) = 3x-9 \\
    y = 3x-9 \\
    x = 3y-9 \\
    x-3y=-9 \\
    x+9=3y \\
    \frac{x+9}{3}=y \\
    f^{-1}(x)=\frac{x+9}{3}
  \end{gathered}
\end{equation}

Domain and range of inverse function:
Domain of a function $f(x)$ is the range of its inverse function $f^{-1}(x)$.
Range of a function $f(x)$ is the domain of its inverse function $f^{-1}(x)$.

\subsection{Multiplicities of the roots of a polynomial}
\begin{equation}
  p(x) = (x-5)^2(x+8)
\end{equation}
for that polynomial roots are $x=5$ and $x=-8$ but factor $(x-5)^2$
appears TWICE so it is a MULTIPLE ROOT
$(x+8)$ on the other hand is a SIMPLE ROOT.

\subsection{Rational functions}
Horizontal asymptotes
1. Identify the 'dominant term' which is a leading term of the polynomial though we
usually ugnore the coefficients.

2. Divide every term in the numerator and denominator of $f(x)$ by the dominant term.

3. Evaluate the remaining expression as $x \to \infty$.

Roots

\subsection{Radical function}
Range of a radical function $f(x) = \sqrt[n]{x}$ depends on index \textit{n} whether it's even or odd.

For odd index like in function $f(x) = -2\sqrt[3]{3x-2}+2$ range is $(-\infty, \infty)$.
Transformations do not change its range.

For even index $n$ range is $x \geqslant 0$ but the transformations can change the range.
Horizontal shifts and stretches doesn't change the the range.

For example for function $g(x) = -3\sqrt[4]{2x-1}-2$:
\begin{equation}
  \begin{gathered}
  \sqrt[4]{2x-1} \geqslant 0 \\
  -3\sqrt[4]{2x-1} \leqslant 0 \\
  -3\sqrt[4]{2x-1} \leqslant -2 \\
  g(x) \leqslant -2
  \end{gathered}
\end{equation}

\subsection{Exponential functions}
\subsection{Logarithmic functions}
\section{Statistics}
\subsection{Linear coefficient}
Signed with $\rho$ (Greek letter called 'rho').
- $\rho = 1$ means perfect positive correlation
- $\rho = -1$ means perfect negative correlation
- $\rho = 0$ no (linear) correlation
\section{Logarithms}
\begin{table}[htbp]
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Property} & \textbf{Formula} \\
\hline
Product Rule & $\log_a(xy) = \log_a(x) + \log_a(y)$ \\
\hline
Quotient Rule & $\log_a\left(\frac{x}{y}\right) = \log_a(x) - \log_a(y)$ \\
\hline
Power Rule & $\log_a(x^n) = n \cdot \log_a(x)$ \\
\hline
Change of Base & $\log_a(x) = \frac{\log_b(x)}{\log_b(a)}$ \\
\hline
Equating arguments & $\log_af(x) = log_ag(x)$ is the same as $f(x) = g(x)$ \\
\hline
\end{tabular}
\caption{Fundamental Properties of Logarithms}
\label{tab:log_properties}
\end{table}

% Approach 1: Looser Table with More Spacing
\begin{table}[htbp]
\centering
\renewcommand{\arraystretch}{1.5} % Increases vertical spacing between rows
\setlength{\tabcolsep}{15pt} % Increases horizontal spacing between columns
\begin{tabular}{|l|p{5cm}|p{6cm}|}
\hline
\textbf{Property} & \textbf{Formula} & \textbf{Description} \\
\hline
\textbf{Product Rule} &
$\log_a(xy) = \log_a(x) + \log_a(y)$ &
The logarithm of a product equals the sum of the individual logarithms. This transforms multiplication into addition. \\
\hline
\textbf{Quotient Rule} &
$\log_a\left(\frac{x}{y}\right) = \log_a(x) - \log_a(y)$ &
The logarithm of a quotient equals the difference of logarithms. This transforms division into subtraction. \\
\hline
\textbf{Power Rule} &
$\log_a(x^n) = n \cdot \log_a(x)$ &
The logarithm of a value raised to a power equals the power multiplied by the logarithm. This transforms exponentiation into multiplication. \\
\hline
\textbf{Change of Base} &
$\log_a(x) = \frac{\log_b(x)}{\log_b(a)}$ &
This formula allows conversion between different logarithmic bases, useful when calculating logarithms with bases other than those built into calculators. \\
\hline
\end{tabular}
\caption{Fundamental Properties of Logarithms}
\label{tab:log_properties}
\end{table}

% Approach 2: List Format with Titles for Each Property
\section*{Logarithmic Properties}

\subsection*{Product Rule}
\begin{equation}
\log_a(xy) = \log_a(x) + \log_a(y)
\end{equation}

The product rule allows us to express the logarithm of a product as the sum of logarithms. This is one of the fundamental properties that makes logarithms powerful for simplifying complex calculations. When we take the logarithm of a product of two numbers, we can break it down into the sum of the logarithms of each number.

\subsection*{Quotient Rule}
\begin{equation}
\log_a\left(\frac{x}{y}\right) = \log_a(x) - \log_a(y)
\end{equation}

The quotient rule states that the logarithm of a quotient equals the logarithm of the numerator minus the logarithm of the denominator. This property transforms division operations into subtraction of logarithms, which is often easier to work with.

\subsection*{Power Rule}
\begin{equation}
\log_a(x^n) = n \cdot \log_a(x)
\end{equation}

The power rule tells us that the logarithm of a number raised to a power equals the power multiplied by the logarithm of the number. This property is especially useful when dealing with expressions containing exponents, as it converts exponentiation to multiplication.

\subsection*{Change of Base Formula}
\begin{equation}
\log_a(x) = \frac{\log_b(x)}{\log_b(a)}
\end{equation}

The change of base formula allows us to convert logarithms from one base to another. This is particularly useful in practical applications since most calculators only directly compute natural logarithms (base $e$) and common logarithms (base 10). Using this formula, we can express logarithms of any base in terms of logarithms of a different base.

\section{Finite sequences}
Coefficient rule:
\begin{equation}
  \sum_{i=1}^{n} ca_i = c * \sum_{i=1}^{n} a_i
\end{equation}

Sum rule:
\begin{equation}
  \sum_{i=1}^{n} (a_i+b_i) = \sum_{i=1}^{n} a_i + \sum_{i=1}^{n} b_i
\end{equation}

\section{Polynomial identities}
\begin{flalign*}
& (a + b)^2 = a^2 + 2ab + b^2 & \\
& (a - b)^2 = a^2 - 2ab + b^2 & \\
& (a + b)^3 = a^3 + 3a^2b + 3ab^2 + b^3 & \\
& (a - b)^3 = a^3 - 3a^2b + 3ab^2 - b^3 & \\
& a^2 - b^2 = (a + b)(a - b) & \\
& a^3 - b^3 = (a - b)(a^2 + ab + b^2) & \\
& a^3 + b^3 = (a + b)(a^2 - ab + b^2) &
\end{flalign*}

\section{Derivatives}
Average rate of change of the function $f(x)$ on the interval $[a, a+h]$

\begin{equation}
  \frac{\Delta y}{\Delta x} = \frac{f(a+h) - f(a)}{h}
\end{equation}

Instanteous rate of change of the function at the point $x=a$ is the limit of the average
rate of change as $h \rightarrow 0$. It's denoted by $f'(a)$ and given by:

\begin{equation}
  f'(a) = \lim_{h \rightarrow 0}\frac{f(a+h) - f(a)}{h}
\end{equation}

and its value is equal to the slope of the tangent line to the curve $y = f(x)$ at $x=a$.
Also it can be defined as:
\begin{equation}
  f'(a) = \lim_{x \rightarrow a}\frac{f(x) - f(a)}{x-a}
\end{equation}

\section{Integrals}
Also called \textit{antiderivatives}. To represent integrals

Integral of a sum of functions
\begin{equation}
  \int f(x) \pm g(x) \, dx = \int f(x) \, dx \pm \int g(x) \, dx
\end{equation}

Integral of a reciprocal function
\begin{equation}
  \int \frac{1}{x} \, dx = \int \ln abs(x) + C
\end{equation}
and also:
\begin{equation}
  \int \frac{1}{x} \, dx = \int \ln|x| + C = \ln abs(x) + \ln K = \ln (K (abs(x))
\end{equation}

Integral of an exponential function
\begin{equation}
  \int e^x \, dx = e^x + C
\end{equation}
with general base we know that
\begin{equation}
  \frac{d}{dx} (a^x) = a^x \ln a
\end{equation}
so reversing:
\begin{equation}
  \int a^x \, dx = \frac{a^x}{\ln a} + C
\end{equation}

Marked integral for future
\begin{equation}
  \int_{a}^{b} f(x) \, dx
\end{equation}

\end{document}