### 2025-08-27 cards/paper.tex
upload bunch of topics for future redaction

\documentclass{article}

\usepackage{amsmath,amssymb,amsthm,textcomp,mathtools}

\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}

\title{On Math}
\author{Jacek Walczak}
\date{February - March 2025}

\begin{document}

\maketitle

\section{General notes and todos}
- I feel like multiline equations should have higher vertical spacing
- Trigonometry values can be done in a single table

\section{Default symbols}

A - area

\section{Trigonometry}
-- TODO: properties of a trigonometric functions
-- like limits, asymptotes etc
\subsection{Properties of a sine function}
zeros of the function
x=n PI
maxima
x = pi/2 + 2npi
minima
x = -pi/2 + 2npi

sine function is odd (which means it has a rotational symmetry around the origin)
sin(-x) = -sin(x)

\subsection{Properties of a cosine function}
\subsection{Properties of a tangent function}
\subsection{Properties of a cosecant function}
y = cscx = 1/sinx
asymptotes: x = n pi
takes a value of 1 at x = pi/2 2npi
takes a value of -1 at x = -pi/2 + 2npi
cosecant is an odd function
\subsection{Properties of a secant function}
y = secx = 1/cosx
asymptotes: x = +- pi/2, +- 3pi/2 etc
takes a value of 1 at x = 2npi
takes a value of -1 at x = pi + 2npi
\subsection{Properties of a cotangent function}
y = cotx = 1 /tanx = cosx/sinx
for zeros x = pi/2 + pi n
vertical asymptotes x = pi n
period pi
cotx is an odd function

\subsection{Pythagorean identity}
\begin{equation}
  \cos^2(\theta) + \sin^2(\theta) = 1
\end{equation}

\begin{equation}
  \begin{gathered}
    \sin\theta = \frac{\text{opposite}}{\text{hypotenuse}} \\
    \cos\theta = \frac{\text{adjacent}}{\text{hypotenuse}} \\
    \tan\theta = \frac{\text{opposite}}{\text{adjacent}} \\
    \csc\theta = \frac{1}{\sin{\theta}} =
    \frac{\text{hypotenuse}}{\text{opposite}} \\
    \sec\theta = \frac{1}{\cos{\theta}} =
    \frac{\text{hypotenuse}}{\text{adjacent}} \\
    \cot\theta = \frac{1}{\tan{\theta}} =
    \frac{\text{adjacent}}{\text{opposite}} \\
  \end{gathered}
\end{equation}

\subsection{Sine values}
\begin{equation}
  \begin{gathered}
    \sin(0) = 0 \\
    \sin(30) = \frac{1}{2} \\
    \sin(45) = \frac{\sqrt{2}}{2} \\
    \sin(60) = \frac{\sqrt{3}}{2} \\
    \sin(90) = 1
  \end{gathered}
\end{equation}

\subsection{Cosine values}
\begin{equation}
  \begin{gathered}
    \cos(0) = 1 \\
    \cos(30) = \frac{\sqrt{3}}{2} \\
    \cos(45) = \frac{\sqrt{2}}{2} \\
    \cos(60) = \frac{1}{2} \\
    \cos(90) = 0
  \end{gathered}
\end{equation}

% Tangent values

\subsection{Tangent values}
\begin{equation}
  \begin{gathered}
    \tan(0) = 0 \\
    \tan(30) = \frac{1}{\sqrt{3}} = \frac{\sqrt{3}}{3} \\
    \tan(45) = 1 \\
    \tan(60) = \sqrt{3} \\
    \tan(90) = \text{undefined}
  \end{gathered}
\end{equation}

\subsection{Formula of a general triangle}

\begin{equation}
  A = \frac{1}{2}ab \sin{C}
\end{equation}
where \textit{C} is the measure of the angle between sides \textit{a} and \textit{b} (also called \textit{included angle})

\subsection{Law of sines}
% TODO: image needed

\begin{equation}
  \frac{a}{\sin A} = \frac{b}{\sin B} = \frac{c}{\sin C}
\end{equation}
and also:
\begin{equation}
  \frac{\sin A}{a} = \frac{\sin B}{b} = \frac{\sin C}{c}
\end{equation}
When we have obtuse angle we need to subtract the result from $180$.

\subsection{Law of cosines}
% TODO: image needed
\begin{equation}
  \begin{gathered}
    a^2 = b^2 + c^2 - 2bc\cos{A} \\
    b^2 = a^2 + c^2 - 2ac\cos{B} \\
    c^2 = a^2 + b^2 - 2ab\cos{C}
  \end{gathered}
\end{equation}

\subsection{Limits in trigonometric functions}
% TODO: image needed
For sine function:
\begin{equation}
  \lim\limits_{x \to a}\sin x = \sin a \\
  \text{and}
  \lim\limits_{x \to \pm\infty}\sin x = DNE
\end{equation}

Similarily for cosine function:
\begin{equation}
  \lim\limits_{x \to a}\cos x = \cos a \\
  \text{and}
  \lim\limits_{x \to \pm\infty}\cos x = DNE
\end{equation}

Differently for tangens though:
\begin{equation}
  \lim\limits_{x \to a}\tan x = \tan a, a \neq \frac{\pi}{2} + n\pi \\
  \text{because there function does not exist! but they are infinite at the sides of a vertical asymptote:} \\
  \lim\limits_{x \to \frac{\pi}{2}^-}\tan x = \infty \\
  \lim\limits_{x \to \frac{\pi}{2}^+}\tan x = -\infty
\end{equation}

\subsection{Double angle formula for sine}
%TODO
sin(x+y) = sinx cosy + cosx siny
substituting x = y
sin2x = 2sinx cosx
it has many applications, it allows to simplyifying some trigonometric expressions
and find the exact value for some non-special angles
this is an identity! not equation, meaning it's true for all xes
\subsection{Double angle formula for cosine}
%TODO

\section{Complex numbers}
Complex numbers can be represented by the point $(x, y)$ where $x$ is real axis and $y$
is imaginary axis. We call this plane as a \textbf{complex plane}, also called
\textit{Argand diagram}, denoted as $C$.

\subsection{Powers of $i$}
\begin{equation}
  \begin{gathered}
    i^1 = \sqrt{-1} \\
    i^2 = (\sqrt{-1})^2 = -1 \\
    i^3 = -1 * i = -i \\
    i^4 = (-1) * (-1) = 1
  \end{gathered}
\end{equation}
and so on...

\subsection{Magnitude of a complex number}
The magnitude of a complex number is its distance from the origin $(0, 0)$.
It is denoted az $|z|$. For complex number $z = a + bi$ magnitude is:
\begin{equation}
  |z| = \sqrt{a^2 + b^2}
\end{equation}

\subsection{Argument of a complex number}
There is natural angle $\theta$ that \textit{z} makes with real x-axis. This angle is called \textit{the argument of z}. The numeric value of $arg(z)$ is given in radians and also
\begin{equation}
  0 \leq arg(z) \leq 2\pi
\end{equation}
Example:
\begin{equation}
  \begin{gathered}
    z = 3 + 4i \\
    \tan \theta = \frac{4}{3} \\
    \theta = \arctan\frac{4}{3} \approx 0.93
  \end{gathered}
\end{equation}

\section{Geometry}
\subsection{Rigid motions}
Or sometimes - any transformations that does not change the distances
and angles between points in a figure. They are \textbf{angle-preserving} and
\textbf{distance-preserving}.

Rigid motions types:

- translations

- rotations

- reflections

Not a rigid motions:

- dilations

- stretches

\subsection{Euler on polyhedrons}
A polyhedron is \textbf{convex} if and only if the following condition
holds:
If we take any two distinct points A and B that belong to the polyhedron
then all the points on the line segment AB also belong to the polyhedron.

Euler's formula for polyhedra for any convex polyhedron, the following
relationship must be satisfied.
\begin{equation}
  V - E + F = 2
\end{equation}

$V$ - number of vertices

$E$ - number of edges

$F$ - number of faces

The \textbf{Euler characteristic} of any polyhedra is denoted by the
Greek letter $\aleph$ (``k-eye'') and is equal to the number:
\begin{equation}
  \aleph = V - E + F
\end{equation}
Any convex polyhedron has
\begin{equation}
  \aleph = 2
\end{equation}
\section{Vectors}
How to call that section?
\subsection{Dot product}
for vectors a, b and c and scalar lambda

- the dot product distributes over addition:
(a+b) * c = a*c + b*c
and
c*(a+b) = a*c + b*c

- the dot product is associative with respect to scalar multiplication
(La) * b = L(a*b)
and
a*(Lb) = L(a*b)

- the dot product is commutative
a*b = b*a

- two non-zero vectors a and b are perpendicular if the angle between them is $\theta = 90$
a perpendicular symbol b

in generat the dot product of two perpendicular vectors is 0. if a perpendicular b and theta is 90 then
a*b = |a| * |b| * cos90 = |a| * |b| * 0 = 9

if a and b are non-zero vectors you can do a conversion
a perp b <=> a*b = 0

\section{Displacement}
\begin{equation}
  s = ut+\frac{1}{2}gt^2
\end{equation}
where:

$s$ - displacement of the object measured in meters

$u$ - initial velocity of the object measured in meters per second

$t$ - time measured in seconds

$g$ - $9.8\frac{m}{s^2}$ - constant acceleration due to gravity
\section{Functions}
\subsection{Standard form}
\begin{equation}
  ax + by = c
\end{equation}
\subsection{Parallel lines}
Calculate equation of the parallel line:
\begin{equation}
  y - y_1 = m(x - x_1)
\end{equation}
Calculate a slope of a line given two points:
\begin{equation}
  m = \frac{y_2 - y_1}{x_2 - x_1}
\end{equation}
\subsection{Quadratic formula}
\begin{equation}
  x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}
\end{equation}

\subsection{Odd and even functions}
Function is said to be odd if:
\begin{equation}
  f(-x) = -f(x)
\end{equation}
Function is said to be even if:
\begin{equation}
  f(-x) = f(x)
\end{equation}
\subsection{Factor cubics by grouping}
For example to factor following function by grouping:

\begin{equation}
  3x^3-6x^2+2x-4
\end{equation}
1. group them
\begin{equation}
  (3x^3-6x^2)+(2x-4)
\end{equation}
2. factor larger terms so the leftover match
\begin{equation}
  3x^2(x-2)+2(x-2)
\end{equation}
3. factor out the binomial
\begin{equation}
  (3x^2+2)(x-2)
\end{equation}
Note: not all cubics can be factored, however,
of the cubics that can be factored,
many can be factored by grouping

\subsection{Factor theorem}
For polynomial, if we know root, we can determine a factor of the
polynomial. If $p(x)$ is a polynomial and $x=r$ is a root of $p(x)$ then
\begin{equation}
  p(r) = 0
\end{equation}
so $(x-r)$ is a factor of $p(x)$
generally: $(ax-b)$ is a factor of a polynomial $p(x)$ then:
\begin{equation}
  x = \frac{b}{a} is a root of p(x)
\end{equation}

\subsection{Rational roots theorem}
Support the rational number $\frac{p}{q}$ in lowest terms, is a root of a polynomial
with integer coefficient.
% TODO: add colors to differentiate
Then 'p' must be the factor of the polynomial's constant term, and 'q' must be a factor
of its leading coefficient.
Given a polynomial with integer coefficient, any rational roots of the polynomial must take the
form:
\begin{equation}
  \pm \frac{factor  of  the  constant  term}{factor  of  the  leading  coefficient}
\end{equation}
For example:
\begin{equation}
  f(x) = 8x^3 + 12x^2 + 6x + 2
\end{equation}
so we have 16 possible roots:
\begin{equation}
  \begin{gathered}
    % TODO: strike through
    \pm \frac{1}{1}, \pm \frac{1}{2}, \pm \frac{1}{4}, \pm \frac{1}{8}, \\
    \pm \frac{2}{1}, \pm \frac{2}{2}, \pm \frac{2}{4}, \pm \frac{2}{8}
  \end{gathered}
\end{equation}
Removing the duplicates, we have 10 possible rational roots:
\begin{equation}
  \pm \frac{1}{8}, \pm \frac{1}{4}, \pm \frac{1}{2}, \pm 1, \pm 2
\end{equation}
Note: it is about \textbf{possible} rational roots.

\subsection{Function composition}
\begin{equation}
  (f \circ g)(x) = f(g(x))
\end{equation}
\subsection{Inverse functions}
Inverse functions are denoted by $f^{-1}(x)$.
For $f(x)$ that maps $x=2$ to $y=0$ an inverse function $f^{-1}(x)$ instantiation would be
$f^{-1}(0)=0$.
Generally it is a simple composition:
\begin{equation}
  f^{-1}(f(x)) = x
\end{equation}
so
\begin{equation}
  (f^{-1} \circ f)(x) = x
\end{equation}

Calculate inverse function:
1. swap $x$ and $y$

2. solve for $y$

Example:
\begin{equation}
  \begin{gathered}
    f(x) = 3x-9 \\
    y = 3x-9 \\
    x = 3y-9 \\
    x-3y=-9 \\
    x+9=3y \\
    \frac{x+9}{3}=y \\
    f^{-1}(x)=\frac{x+9}{3}
  \end{gathered}
\end{equation}

Domain of a function $f(x)$ is the range of its inverse function $f^{-1}(x)$.

Range of a function $f(x)$ is the domain of its inverse function $f^{-1}(x)$.

\subsection{Multiplicities of the roots of a polynomial}
\begin{equation}
  p(x) = (x-5)^2(x+8)
\end{equation}
for that polynomial roots are $x=5$ and $x=-8$ but factor $(x-5)^2$
appears \textbf{twice} so it is a \textbf{multiple root}
$(x+8)$ on the other hand is a \textbf{simple root}.

\subsection{Intersection of a reciprocal function and a straight line}
There are three possibilities when considering the intersection of a reciprocal
function and a straight line:
1. there are two intersection points
2. there is only one intersection point
3. there are no intersection points at all

% TODO: needs images

Typically, when computing the points of intersection of a reciprocal
function and a line, we have to solve a system of two equations.

% TODO: needs bracket over two first lines
\begin{equation}
  \begin{gathered}
    y = x \\
    y = \frac{1}{x} \\
    x = \frac{1}{x} \\
    x^2 = 1 \\
    x = \pm 1
  \end{gathered}
\end{equation}

\subsection{Rational functions}
Horizontal asymptotes

1. Identify the 'dominant term' which is a leading term of the polynomial though we
usually ugnore the coefficients.

2. Divide every term in the numerator and denominator of $f(x)$ by the dominant term.

3. Evaluate the remaining expression as $x \to \infty$.

Roots

\subsection{Radical function}
Range of a radical function $f(x) = \sqrt[n]{x}$ depends on index \textit{n} whether it's even or odd.

For odd index like in function $f(x) = -2\sqrt[3]{3x-2}+2$ range is $(-\infty, \infty)$.
Transformations do not change its range.

For even index $n$ range is $x \geqslant 0$ but the transformations can change the range.
Horizontal shifts and stretches doesn't change the the range.

For example for function $g(x) = -3\sqrt[4]{2x-1}-2$:
\begin{equation}
  \begin{gathered}
    \sqrt[4]{2x-1} \geqslant 0 \\
    -3\sqrt[4]{2x-1} \leqslant 0 \\
    -3\sqrt[4]{2x-1} \leqslant -2 \\
    g(x) \leqslant -2
  \end{gathered}
\end{equation}

\subsection{Reciprocal functions}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \begin{axis}[
      axis lines=center,
      xlabel=$x$,
      ylabel=$y$,
      title={Reciprocal Function $f(x) = \frac{1}{x}$},
      xmin=-5, xmax=5,
      ymin=-3, ymax=3,
      xtick={-5,-4,-3,-2,-1,1,2,3,4,5},
      ytick={-3,-2,-1,1,2,3},
      grid=both,
      domain=-5:5,
      samples=200,
      restrict y to domain=-3:3,
      ]
      % Plot the function, excluding the region around x=0
      \addplot[blue, thick] {1/x};

      % Add asymptotes
      \draw[dashed, red] (axis cs:0,-3) -- (axis cs:0,3);
      \draw[dashed, red] (axis cs:-5,0) -- (axis cs:5,0);
    \end{axis}
  \end{tikzpicture}
  \caption{Graph of the reciprocal function $f(x) = \frac{1}{x}$}
\end{figure}

% TODO: image
\begin{equation}
  f(x) = \frac{1}{x}
\end{equation}
When function is not transformed, it passes through points $(-1, -1)$ and $(1, 1)$.
% TODO: not range?
Domain of (transformed) reciprocal function
\begin{equation}
  f(x) = \frac{a}{bx+c} + d
\end{equation}
has the horizontal asymptote $y = d$ therefore the range is
\begin{equation}
  f(x) = \in (-\infty, d) \cup (d, \infty)
\end{equation}
% TODO: range
Range of (transformed) reciprocal function
\begin{equation}
  f(x) = \frac{a}{bx+c} + d
\end{equation}
has the vertical asymptote $y = d$ therefore the range is
\begin{equation}
  f(x) = \in (-\infty, d) \cup (d, \infty)
\end{equation}

\subsection{Exponential functions}
\subsection{Logarithmic functions}
\subsection{Cubic curves functions}
general function
\begin{equation}
  f(x) = ax^3 + bx^2 + cx + d
\end{equation}
range:
\begin{equation}
  y \in (-\infty, \infty)
\end{equation}
At least one real root and at most three distinct roots. For example:
\begin{equation}
  \begin{gathered}
    y = x^3 - x \\
    0 = x(x^2 - 1) \\
    0 = x(x - 1)(x + 1) \\
    x = -1, 0, 1
  \end{gathered}
\end{equation}
%TODO image

- for single roots, the curve passes through the $x-axis$.

- touches the $x-axis$ and turns around at the double root.

- for root of multiplicity 3, the graph becomes \textit{tangent} (lies parallel) to the $x-axis$ and crosses it.

\subsection{Function continuity}
If a function $f(x)$ is \textbf{continuous}, we can draw $y = f(x)$ without
taking our pen off the paper. A function $f(x)$ is continuous at the point
$x = c$ if:

\begin{equation}
  \lim_{x \to c}f(x) = f(c)
\end{equation}

To test for continuity of $f(x)$ we need to check three conditions:

1. does $f(c)$ exists?
We must check that $f(x)$ is defined at $x=c$

2. does $\lim_{x \to c}f(x)$ exists?
We must have to check that $\lim_{x \to c^-}f(x) = \lim{x \to c^+}f(x)$

3. does $\lim{x \to c}f(x) = f(c)$

If all of the above is yes, then the function \textbf{is continuous} at $x = c$.
Otherwise we say there's \textbf{discontinuity} at $x = c$.

\begin{tikzpicture}[scale=1.2]
    % Draw axes
    \draw[->] (-0.5,0) -- (5,0) node[right] {$x$};
    \draw[->] (0,-0.5) -- (0,4) node[above] {$y$};

    % Draw the function curve (looks like a logarithmic or similar curve)
    % Left part approaching the discontinuity
    \draw[thick] plot[domain=0.1:5, samples=100] (\x, {1.5 + 0.8*ln(\x)});

    % Mark the discontinuity point
    \coordinate (c) at (3,0);
    \coordinate (leftlimit) at (3,2.38);

    % Draw vertical dashed line at x = c
    \draw[dashed] (3,0) -- (3,3.5);

    % Draw horizontal dashed line showing limit
    \draw[dashed] (0,2.4) -- (3,2.4);

    % Mark points
    \fill (leftlimit) circle (2pt);

    % Add labels
    \node[below] at (3,0) {$c$};
    \node[left] at (0,2.4) {$\lim_{x \to c}f(x)$};
    \node[right] at (4.5,2.5) {$y = f(x)$};

    % Add limit annotations
    \node at (2,0.5) {$x \to c^-$};
    \node at (4,0.5) {$x \to c^+$};

    % Add arrows showing approach for x
    \draw[->] (2.5,0.3) -- (2.8,0.3);
    \draw[->] (3.5,0.3) -- (3.2,0.3);

    % Add arrows showing approach for y
    \node at (0.5,2.8) {$\downarrow f(x)$};
    \node at (0.5,2.0) {$\uparrow f(x)$};

\end{tikzpicture}

\subsection{Continuity over an interval}
In general it is said that function $f(x)$ is continuous over the interval $[a, b]$ if it is:

- continuous on the open interval $(a, b)$

- right continuous at $x = a$

- left continuous at $x = b$

\section{Probability}
\subsection{Independent events}
In probability, we say that two events are \textbf{independent} when the outcome of one
event doesn't influence the outcome of the other event. Otherwise, the events are \textbf{dependent}.

As for multiplication law for independent events:
\begin{equation}
  P(A \cap B) = P(A) * P(B)
\end{equation}
but, if A and B are independent, then the probability of B occuring is not influenced by
whether or not A has occured, so:
\begin{equation}
  P(B|A) = P(B)
\end{equation}
and replacing it for the first law it goes:
\begin{equation}
  P(A \cap B) = P(A) * P(B)
\end{equation}

\subsection{The addition law of probability}
\begin{equation}
  P(A \cup B) = P(A) + P(B) - P(A \cap B)
\end{equation}
There's some ince intuition behind this law: by adding $P(A)$ to $P(B)$, we count $P(A \cap B)$ twice, hence we need to get rid of one.

\subsection{The multiplication law for conditional probability}
\begin{equation}
  P(A|B) = \frac{P(A \cap B)}{P(B)}
\end{equation}
% TODO provide an example
\subsection{The law of total probability}
% TODO

\section{Series and Sequences}
It is important to understand the difference between a sequence and a series.
A sequence is a list of numbers (separated by commmas).
A series is what you get when you add all of the terms of a sequence together.
\subsection{The sum of an arithmetic series}
\begin{equation}
  S_N = \sum^N_{n=1}a_n
\end{equation}
alternatively
\begin{equation}
  S_N = \frac{N}{2}(a_1 + a_N)\
end{equation}
also, if you don't know the last element of a series, you can use the following:
\begin{equation}
  S_N = \frac{N}{2}(a_1 + a_N) = \frac{N}{2}(2a_1 + d(N-1))
\end{equation}

\section{Statistics}
\subsection{Estimating a mean and variance}
\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|l|}
    \hline
    \textbf{Reading times (in minutes)} & \textbf{Frequency} \\
    \hline
    $0 \leqslant x < 20$ & $5$ \\
    \hline
    $20 \leqslant x < 40$ & $10$ \\
    \hline
    $40 \leqslant x < 60$ & $8$ \\
    \hline
    $60 \leqslant x < 80$ & $7$ \\
    \hline
  \end{tabular}
  \caption{Estimating a mean and variance}
  \label{tab:estimating_a_mean_and_variance}
\end{table}
To estimate a mean we need a midpoint for $a \leqslant x < b$, so midpoint is $\frac{a+b}{2}$
\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|l|l|l|}
    \hline
    \textbf{Reading times (in minutes)} & \textbf{Frequency} & \textbf{Midpoint} & \textbf{F*M} \\
    \hline
    $0 \leqslant x < 20$ & $5$ & $10$ & $50$ \\
    \hline
    $20 \leqslant x < 40$ & $10$ & $30$ & $300$ \\
    \hline
    $40 \leqslant x < 60$ & $8$ & $50$ & $400$ \\
    \hline
    $60 \leqslant x < 80$ & $7$ & $70$ & $490$ \\
    \hline
  \end{tabular}
  \caption{Not sure what to type in here}
  \label{tab:estimating_a_mean_and_variance_2}
\end{table}
\begin{equation}
  \overline{\rm x} = \frac{50+300+400+490}{30} = \frac{1240}{30} \approx 41.3
\end{equation}
Mean estimation for continous grouped data:
\begin{equation}
  \overline{\rm x} \approx \frac{1}{n}\Sigma^K_{i=1}(f_i * m_i)
\end{equation}

$f_i$ - frequency of the i-th class

$m_i$ - midpoint of the i-th class

$n$ - total number of the observations

Variance estimation
\begin{equation}
  \sigma^2_n \approx \frac{1}{n}\Sigma^K_{i=1}(f_i * (m_i - \overline{\rm x})^2)
\end{equation}

$f_i$ - frequency of the i-th class

$m_i$ - midpoint of the i-th class

$n$ - total number of the observations

$\overline{\rm x}$ - the mean

\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|l|l|l|l|}
    \hline
    \textbf{Midpoint} & \textbf{F*M} & \textbf{$M_i * \overline{\rm x}$} & \textbf{($M_i - \overline{\rm x})^2$} & \textbf{$f_i * (M_i - \overline{\rm x})$} \\
    \hline
    $10$ & $50$ & $-31.333$ & $981.776$ & $4908.879$ \\
    \hline
    $10$ & $30$ & $-11.333$ & $128.444$ & $1284.437$ \\
    \hline
    $50$ & $400$ & $8.667$ & $75.112$ & $600.894$ \\
    \hline
    $70$ & $490$ & $28.667$ & $821.780$ & $5752.458$  \\
    \hline
  \end{tabular}
  \caption{Not sure what to type in here}
  \label{tab:estimating_a_mean_and_variance_3}
\end{table}
\begin{equation}
  \begin{gathered}
    \sigma^2_n \approx \frac{4908... + 1284... + 600... + 5752}{30} \approx 418.2  \\
    \sigma_n \approx \sqrt{418.2} \approx 20.5
  \end{gathered}
\end{equation}
\subsection{Sum of squares}
Given that variance is:
\begin{equation}
  \sigma^2_x = \frac{1}{n} \sum_{i=1}^n(x_i - \overline{\rm x})^2
\end{equation}
The sum of squares is the same but with no fraction:
\begin{equation}
  S_{xx} = \sum_{i=1}^n(x_i - \overline{\rm x})^2
\end{equation}
Compare covariance with a sum of squares:
\begin{equation}
  \begin{gathered}
    Cov(x,y) = \frac{1}{n} \sum_{i=1}^n(x_i - \overline{\rm x})(y_i - \overline{\rm y}) \\
    S_{xy} = \sum_{i=1}^n(x_i - \overline{\rm x})(y_i - \overline{\rm y})
  \end{gathered}
\end{equation}
Variance and covariance in terms of $S_{xx}$, $S_{yy}$ and $S_{xy}$:
\begin{equation}
  \begin{gathered}
    \sigma^2_x = \frac{S_{xx}}{n} \\
    \sigma^2_y = \frac{S_{yy}}{n} \\
    Cov(x,y) = \frac{S_{xy}}{n}
  \end{gathered}
\end{equation}
% TODO: some heading?
Shortcut formulas:
\begin{equation}
  \begin{gathered}
    S_{xx} = \sum_{i=1}^n x_i^2 - \frac{1}{n} (\sum_{i=1}^n x_i)^2 \\
    S_{yy} = \sum_{i=1}^n y_i^2 - \frac{1}{n} (\sum_{i=1}^n y_i)^2 \\
    S_{xy} = \sum_{i=1}^n x_iy_i - \frac{1}{n} (\sum_{i=1}^n x_i)(\sum_{i=1}^n y_i)
  \end{gathered}
\end{equation}
\subsection{Linear regression}
For a set of $n$ paired observations $x_i$ and $y_i$ the trend line is the
straight line that best fits the data.
Finding the equation of the trend line is called \textbf{linear regression}
or best fit.
We can show that the equation of the trend line is given by:
\begin{equation}
  y = mx * b
\end{equation}
1. Slope $m$
\begin{equation}
  m = \frac{S_{xy}}{S_{xx}}
\end{equation}
2. y-intercept - b:
\begin{equation}
  b = \overline{\rm y} - m\overline{\rm x}
\end{equation}
3. $\overline{\rm x}$ and $\overline{\rm y}$ are the means of the data sets:
\begin{equation}
  \overline{\rm x} = \frac{1}{n} \sum_{i=1}^nx_i, \overline{\rm y} = \frac{1}{n} \sum_{i=1}^ny_i,
\end{equation}
4. quantities $S_{xx}$ and $S_{xy}$:
\begin{equation}
  \begin{gathered}
    S_{xx} = \sum_{i=1}^n (x_i - \overline{\rm x})^2 \\
    S_{xy} = \sum_{i=1}^n(x_i - \overline{\rm x})(y_i - \overline{\rm y})
  \end{gathered}
\end{equation}
There is some nice intuition behind the formula for $b$. The trend line
always passes through the point $(\overline{\rm x}, \overline{\rm y})$.
Therefore, this point must satisfy the trend line equation:
\begin{equation}
  y = mx + b
\end{equation}
We can then solve this equation for $b$. Also notice that $S_{xx}$ and $S_{xy}$
are the summation terms that feature in the formulas for the variance and
covariance respectively.
\subsection{Linear coefficient}
Signed with $\rho$ (Greek letter called 'rho').
- $\rho = 1$ means perfect positive correlation
- $\rho = -1$ means perfect negative correlation
- $\rho = 0$ no (linear) correlation
\subsection{Linear correlation coefficient}
$|p| > 0.7$ - correlation is strong
$0.3 \leqslant |p| \leqslant 0.7$ - correlation is weak
$0 \leqslant |p| \leqslant 0.3$ - correlation is negligible
\subsection{Linear correlation coefficient using sigma notation}
\begin{equation}
  p = \frac{S_{xy}}{\sqrt{S_{xx} * S_{yy}}}
\end{equation}
\subsection{Residuals and residual plots}
For dataset of paired numerical observations $(x, y)$ and the corresponding
linear regression model:
\begin{equation}
  \hat y = f(x)
\end{equation}
The \textbf{residual} for a particular observation $(x, y)$ is defined as:
\begin{equation}
  \begin{gathered}
    \text{residual} = \text{actual} - \text{estimated} \\
    y - f(x) = y - \hat y
  \end{gathered}
\end{equation}
We subtract the estimated value predicted by model from the actual observed
value:
\begin{itemize}
\item negative residual - model overestimates the true value
\item positive residual - model underestimates the true value
\end{itemize}

Exercises:
\begin{equation}
  \begin{gathered}
    \hat y = 0.02x + 0.8 \\
    \text{for point} P(400, 8.3) \\
    \hat y = 0.02 * 400 + 0.8 = 8.8 \\
    y - \hat y = 8.3 - 8.8 = -0.5
  \end{gathered}
\end{equation}
\section{Logarithms}
\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|l|}
    \hline
    \textbf{Property} & \textbf{Formula} \\
    \hline
    Product Rule & $\log_a(xy) = \log_a(x) + \log_a(y)$ \\
    \hline
    Quotient Rule & $\log_a\left(\frac{x}{y}\right) = \log_a(x) - \log_a(y)$ \\
    \hline
    Power Rule & $\log_a(x^n) = n \cdot \log_a(x)$ \\
    \hline
    Change of Base & $\log_a(x) = \frac{\log_b(x)}{\log_b(a)}$ \\
    \hline
    Equating arguments & $\log_af(x) = log_ag(x)$ is the same as $f(x) = g(x)$ \\
    \hline
  \end{tabular}
  \caption{Fundamental Properties of Logarithms}
  \label{tab:log_properties}
\end{table}

% Approach 1: Looser Table with More Spacing
\begin{table}[htbp]
  \centering
  \renewcommand{\arraystretch}{1.5} % Increases vertical spacing between rows
  \setlength{\tabcolsep}{15pt} % Increases horizontal spacing between columns
  \begin{tabular}{|l|p{5cm}|p{6cm}|}
    \hline
    \textbf{Property} & \textbf{Formula} & \textbf{Description} \\
    \hline
    \textbf{Product Rule} &
                            $\log_a(xy) = \log_a(x) + \log_a(y)$ &
                                                                   The logarithm of a product equals the sum of the individual logarithms. This transforms multiplication into addition. \\
    \hline
    \textbf{Quotient Rule} &
                             $\log_a\left(\frac{x}{y}\right) = \log_a(x) - \log_a(y)$ &
                                                                                        The logarithm of a quotient equals the difference of logarithms. This transforms division into subtraction. \\
    \hline
    \textbf{Power Rule} &
                          $\log_a(x^n) = n \cdot \log_a(x)$ &
                                                              The logarithm of a value raised to a power equals the power multiplied by the logarithm. This transforms exponentiation into multiplication. \\
    \hline
    \textbf{Change of Base} &
                              $\log_a(x) = \frac{\log_b(x)}{\log_b(a)}$ &
                                                                          This formula allows conversion between different logarithmic bases, useful when calculating logarithms with bases other than those built into calculators. \\
    \hline
  \end{tabular}
  \caption{Fundamental Properties of Logarithms}
  \label{tab:log_properties}
\end{table}

% Approach 2: List Format with Titles for Each Property
\section*{Logarithmic Properties}

\subsection*{Product Rule}
\begin{equation}
  \log_a(xy) = \log_a(x) + \log_a(y)
\end{equation}

The product rule allows us to express the logarithm of a product as the sum of logarithms. This is one of the fundamental properties that makes logarithms powerful for simplifying complex calculations. When we take the logarithm of a product of two numbers, we can break it down into the sum of the logarithms of each number.

\subsection*{Quotient Rule}
\begin{equation}
  \log_a\left(\frac{x}{y}\right) = \log_a(x) - \log_a(y)
\end{equation}

The quotient rule states that the logarithm of a quotient equals the logarithm of the numerator minus the logarithm of the denominator. This property transforms division operations into subtraction of logarithms, which is often easier to work with.

\subsection*{Power Rule}
\begin{equation}
  \log_a(x^n) = n \cdot \log_a(x)
\end{equation}

The power rule tells us that the logarithm of a number raised to a power equals the power multiplied by the logarithm of the number. This property is especially useful when dealing with expressions containing exponents, as it converts exponentiation to multiplication.

\subsection*{Change of Base Formula}
\begin{equation}
  \log_a(x) = \frac{\log_b(x)}{\log_b(a)}
\end{equation}

The change of base formula allows us to convert logarithms from one base to another. This is particularly useful in practical applications since most calculators only directly compute natural logarithms (base $e$) and common logarithms (base 10). Using this formula, we can express logarithms of any base in terms of logarithms of a different base.

\section{Finite sequences}
Coefficient rule:
\begin{equation}
  \sum_{i=1}^{n} ca_i = c * \sum_{i=1}^{n} a_i
\end{equation}

Sum rule:
\begin{equation}
  \sum_{i=1}^{n} (a_i+b_i) = \sum_{i=1}^{n} a_i + \sum_{i=1}^{n} b_i
\end{equation}

\section{Polynomials}
\subsection{Polynomial identities}
\begin{flalign*}
  & (a + b)^2 = a^2 + 2ab + b^2 & \\
  & (a - b)^2 = a^2 - 2ab + b^2 & \\
  & (a + b)^3 = a^3 + 3a^2b + 3ab^2 + b^3 & \\
  & (a - b)^3 = a^3 - 3a^2b + 3ab^2 - b^3 & \\
  & a^2 - b^2 = (a + b)(a - b) & \\
  & a^3 - b^3 = (a - b)(a^2 + ab + b^2) & \\
  & a^3 + b^3 = (a + b)(a^2 - ab + b^2) &
\end{flalign*}
\subsection{End behavior of general polynomials}
Given polynomial
\begin{equation}
  f(x) = a_nx^n + a_{n-1}x^{n-1} + ... +a_1x + a_0
\end{equation}
with leading coefficient $a \neq 0$ the end behavior is always
\begin{equation}
  f(x) \rightarrow \infty
\end{equation}
or
\begin{equation}
  f(x) \rightarrow -\infty
\end{equation}
as
\begin{equation}
  x \rightarrow \pm \infty
\end{equation}
If the leading coefficient is positive then
$f(x) \rightarrow \infty$ as $x \rightarrow \infty$,
if negative then
$f(x) \rightarrow -\infty$ as $x \rightarrow \infty$
If leading exponent is \textbf{even}, then the end behavior as $x \rightarrow \infty$ \textbf{is the same} as the end behavior as $x \rightarrow -\infty$.
Otherwise, if the leading exponent is \textbf{odd}, then the end behavior as $x \rightarrow +\infty$ is the opposite as $x \rightarrow -\infty$.
\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|l|l|l|}
    \hline
    \textbf{$n$} & \textbf{$a_n$} & \textbf{$x \rightarrow -\infty$} & \textbf{$x \rightarrow +\infty$} \\
    \hline
    even & $>0$ & $f(x) \rightarrow +\infty$ & $f(x) \rightarrow +\infty$ \\
    \hline
    even & $<0$ & $f(x) \rightarrow -\infty$ & $f(x) \rightarrow -\infty$ \\
    \hline
    odd & $>0$ & $f(x) \rightarrow -\infty$ & $f(x) \rightarrow +\infty$ \\
    \hline
    odd & $<0$ & $f(x) \rightarrow +\infty$ & $f(x) \rightarrow -\infty$ \\
    \hline
  \end{tabular}
  \caption{End behavior of polynomials}
  \label{tab:end_behavior_of_polynomials}
\end{table}
Mnemonic trick:
substitute $+1$ and $-1$ into the leading term of polynomial, for example:
\begin{equation}
  \begin{gathered}
    P(x) = 2x^3 - 6x^2 + 8 \\
    \\
    2(+1)^3 = +2 \\
    P(x) \rightarrow +\infty \\
    \text{as} \\
    x \rightarrow +\infty \\
    \\
    2(-1)^3 = -2 \\
    P(x) \rightarrow -\infty \\
    \text{as} \\
    x \rightarrow -\infty \\
  \end{gathered}
\end{equation}
In general whenever $n$ is an \textit{even natural number} we have
\begin{equation}
  \begin{gathered}
    \lim_{x \to +\infty}x^n = +\infty \\
    \text{and} \\
    \lim_{x \to -\infty}x^n = +\infty
  \end{gathered}
\end{equation}
for $n = 2, 4, 6, ...$ but for \textit{odd natural number} though:
\begin{equation}
  \begin{gathered}
    \lim_{x \to +\infty}x^n = +\infty \\
    \text{and} \\
    \lim_{x \to -\infty}x^n = -\infty
  \end{gathered}
\end{equation}
for $n = 1, 3, 5, ...$. To compute the limit at infinity of general polynomial, we only need
to consider the term with the largest expontent, called the \textbf{leading term}.
All of the other terms can be disregarded.
\begin{equation}
  \begin{gathered}
    \lim_{x \to \infty}(2x^3 - 7x^2 + 6x + 12) = \\
    = \lim_{x \to \infty}(2x^3) = \\
    = \lim_{x \to \infty}(x^3) = \\
    = \infty
  \end{gathered}
\end{equation}
\section{Derivatives}
Average rate of change of the function $f(x)$ on the interval $[a, a+h]$:
\begin{equation}
  \frac{\Delta y}{\Delta x} = \frac{f(a+h) - f(a)}{h}
\end{equation}
Instanteous rate of change of the function at the point $x=a$ is the limit of the average
rate of change as $h \rightarrow 0$. It's denoted by $f'(a)$ and given by:
\begin{equation}
  f'(a) = \lim_{h \rightarrow 0}\frac{f(a+h) - f(a)}{h}
\end{equation}
and its value is equal to the slope of the tangent line to the curve $y = f(x)$ at $x=a$.
Also it can be defined as:
\begin{equation}
  f'(a) = \lim_{x \rightarrow a}\frac{f(x) - f(a)}{x-a}
\end{equation}

\subsection{Tangent line equation}
% TODO image
Equation of the tangent line to the curve $y=f(x)$ at the point $P(x_1, y_1)$ is given by
\begin{equation}
  \begin{gathered}
    y - y_1 = m(x - x_1) \\
    \text{where:} \\
    m = f'(x_1)
  \end{gathered}
\end{equation}

\subsection{Normal line equation}
% TODO image
Equation of the normal line to the curve $y=f(x)$ at the point $P(x_1, y_1)$ is given by
\begin{equation}
  \begin{gathered}
    y - y_1 = m(x - x_1) \\
    \text{where:} \\
    m = -\frac{1}{f'(x_1)}
  \end{gathered}
\end{equation}
Normal line is just negative reciprocal of the slope of the \textbf{tangent line}.

\subsection{Chain rule for differentiation}
Derivative of a composite function $f(g(x))$ is given by:
\begin{equation}
  \frac{df}{dx} = \frac{df}{dg} * \frac{dg}{dx} = \frac{df}{dx}
\end{equation}
Example:
\begin{equation}
  \begin{gathered}
    f(x) = (3x^4 + x^3)^5 \\
    \frac{df}{dx} = ? \\
    f = g^5 \\
    g = 3x^4 + x^3 \\
    \frac{df}{dx} = \frac{df}{dg} * \frac{dg}{dx} \\
    \frac{df}{dg} = 5g^4 \\
    \frac{dg}{dx} = 12x^3 + 3x^2 \\
    \frac{df}{dx} = 5g^4 * (12x^3 + 3x^2) = 5(3x^4 + x^3)^4(12x^3 + 3x^2)
  \end{gathered}
\end{equation}
The chain rule also can be expressed using primes instead of Leibniz notation
\begin{equation}
  f(g(x))' = f'(g(x)) * g'(x)
\end{equation}
\subsection{Product rule for differentiation}
for $f(x) = u(x)*v(x)$ differentiation is:
\begin{equation}
  \begin{gathered}
    f'(x) = u'(x)*v(x) + u(x)*v'(x) \\
    \text{Using Leibnitz:} \\
    \frac{dy}{dx} = \frac{du}{dx}*v + \frac{dv}{dx}*u
  \end{gathered}
\end{equation}
\subsection{Quotient rule for differentiation}
for $f(x) = \frac{u(x)}{v(x)}$ differentiation is:
\begin{equation}
  \begin{gathered}
    f'(x) = \frac{u'(x)*v(x) - u(x)*v'(x)}{[v(x)]^2} \\
    \text{Using Leibnitz:} \\
    \frac{d}{dx}(\frac{u}{v}) = \frac{v*\frac{du}{dx} - u*\frac{dv}{dx}}{v^2}
  \end{gathered}
\end{equation}
\subsection{Differentiation of an exponent function}
\begin{equation}
  \frac{d}{dx} (a^x) = a^x\ln a
\end{equation}
with special case:
\begin{equation}
  \frac{d}{dx} (e^x) = e^x\ln e = e^x * 1 = e^x
\end{equation}
\subsection{Chain rule for exponent function}
\begin{equation}
  \begin{gathered}
    y = e^{3x} \\
    y = e^u \\
    u = 3x \\
    \frac{dy}{dx} = \frac{dy}{du} * \frac{du}{dx} \\
    \frac{d}{du} e^u = e^u \\
    \frac{d}{dx} 3x = 3 \\
    \frac{dy}{dx} = e^u * 3 = 3e^u = 3e^{3x}
  \end{gathered}
\end{equation}

\subsection{Derivative of the (natural) logarithm}
For natural logarithm:
\begin{equation}
  \frac{d}{dx}(\ln x) = \frac{1}{x}
\end{equation}
For general logarithm:
\begin{equation}
  \frac{d}{dx}(\log_a x) = \frac{1}{x\ln a}
\end{equation}
Example with changing base:
\begin{equation}
  \begin{gathered}
    \log_ax = \frac{\ln x}{\ln a} \\
    \frac{d}{dx}(\log_ax) = \frac{d}{dx}(\frac{\ln x}{\ln a}) = \\
    = \frac{1}{\ln a} * \frac{d}{dx}(\ln x) = \\
    = \frac{1}{\ln a} * \frac{1}{x} = \frac{1}{x\ln a}
  \end{gathered}
\end{equation}
\subsection{Chain rule for logarithmic function}
\begin{equation}
  \begin{gathered}
    y = \ln(2x+1) \\
    y = \ln u \\
    u = 2x+1 \\
    y' = \frac{1}{u} \\
    u' = 2 \\
    \frac{dy}{dx} = \frac{1}{u} * 2 = \frac{2}{u} = \frac{2}{2x+1}
  \end{gathered}
\end{equation}

\subsection{Second and higher order derivatives}
The second order derivative is the derivative of a functions first derivative.
It can be written as $f''(x)$. Also
\begin{equation}
  \begin{gathered}
    f''(x) = (f'(x))' \\
    \text{example:} \\
    f(x) = 4x^2 + 3x \\
    f'(x) = 8x+3 \\
    f''(x) = 8
  \end{gathered}
\end{equation}
For Leibnitz notation it can be written as:
\begin{equation}
  \begin{gathered}
    f''(x) = \frac{d^2f}{dx^2} = \frac{d}{dx}(\frac{df}{dx}) \\
    \text{or, for y:} \\
    \frac{d^2y}{dx^2} \\
    \text{for higher order, in general:} \\
    f^{(n)}(x) = \frac{d^n}{dx^n}f(x)
  \end{gathered}
\end{equation}
\section{Integrals}
Also called \textit{antiderivatives}. To represent integrals we use an integral
symbol. For example:
\begin{equation}
  \int 2x dx = x^2 + C
\end{equation}
Integrals like this are called 'indefinite integrals' and (in this case)
function being integrated is called an 'integrand'.

\subsection{Integral of a sum of functions}
\begin{equation}
  \int f(x) \pm g(x) \, dx = \int f(x) \, dx \pm \int g(x) \, dx
\end{equation}

\subsection{Integral of a reciprocal function}
\begin{equation}
  \int \frac{1}{x} \, dx = \int \ln abs(x) + C
\end{equation}
and also:
\begin{equation}
  \int \frac{1}{x} \, dx = \int \ln|x| + C = \ln abs(x) + \ln K = \ln (K (abs(x))
\end{equation}
% TODO
more accurate I guess:
\begin{equation}
  \int \frac{1}{ax+b} \, dx = \frac{1}{a} \ln|ax+b| + C
\end{equation}

\subsection{Integral of an exponential function}
\begin{equation}
  \int e^x \, dx = e^x + C
\end{equation}
with general base we know that
\begin{equation}
  \frac{d}{dx} (a^x) = a^x \ln a
\end{equation}
so reversing:
\begin{equation}
  \int a^x \, dx = \frac{a^x}{\ln a} + C
\end{equation}

\subsection{Integrals of trigonometry functions}
\begin{equation}
  \begin{gathered}
    \int \sin x dx = -\cos x + C \\
    \int \cos x dx = \sin x + C \\
    \int \sec^2 x dx = \tan x + C \\
    \int \sec x \tan x dx = \sec x + C \\
    \int \csc x \cot x dx = -\csc x + C \\
    \int \csc^2 x dx = -\cot x + C
  \end{gathered}
\end{equation}

\subsection{Power rule for differentiation in reverse}
\begin{equation}
  \frac{d}{dx}(x^{n+1}) = (n+1)x^n
\end{equation}
same as:
\begin{equation}
  \int x^ndx = \frac{x^{n+1}}{n+1}
\end{equation}
Note: power rule does not work if $n=-1$!

\subsection{Left Riemann sum}
- underestimates the actual area under the curve, if the function is increasing

- overestimates, if the function is decreasing

\subsection{Implicit differentiation}
%TODO is it a good place?

\subsection{Integrating with substitution}
%TODO example \int 8(2x+5)^3dx
general formula
\begin{equation}
  \int(ax+b)^n dx = \frac{1}{a(n+1)}(ax+b)^{n+1} + C, n \neq -1
\end{equation}

\section{Sandbox}
Marked integral for future
\begin{equation}
  \int_{a}^{b} f(x) \, dx
\end{equation}

\end{document}